---
title: "Plotting temporal changes to exploitable fish biomass in the Southern Ocean"
author: "Denisse Fierro Arcos"
date: "2024-02-19"
format: gfm
toc: true
editor: visual
---

# Changes in total consumer biomass in the Southern Ocean over time

In this notebook, we will use all FishMIP global models to calculate the mean ensemble percentage change in total exploitable fish biomass in the Southern Ocean for the decade ending in 2050. The reference period for this calculation is 2005-2014. We define "exploitable fish biomass" as biomass for all marine organisms with weights between 10g and 100Kg.  
  
## Loading libraries

```{r}
#| warning: false
#| output: false

#Data wrangling
library(tidyverse)
library(data.table)
#Combining plots
library(cowplot)
```

## Setting up notebook
We will define the folders where inputs are kept, and where outputs should be saved.  
  
```{r}
#Base folder for project
base_folder <- "/rd/gem/private/users/camillan"

#Defining location of notebook outputs
out_folder <- "../outputs"
if(!dir.exists(out_folder)){
  dir.create(out_folder)
}
```
  
## Loading data
We will load the total consumer biomass data to calculate percentage change, and masks to extract data for the MEASO and CCAMLR regions only.  
  
```{r}
#MEASO
measo_mask <- read_csv("../outputs/measo_regions_1deg.csv")
measo_keys <- read_csv("../outputs/measo_regions_keys.csv")

#CCAMLR
ccamlr_mask <- read_csv("../outputs/ccamlr_stats_area_1deg.csv")
ccamlr_keys <- read_csv("../outputs/ccamlr_stats_area_keys.csv")

#Getting a list of files containing biomass data
global_files <- list.files(base_folder, pattern = "global_10g-100kg.rds", 
                           full.names = T, recursive = T)

#Getting a list of models
members <- str_extract(global_files, "annual_(.*)_(h|s)", group = 1) |> 
  unique()
```
  
## Calculating percentage change in fish biomass estimates from global FishMIP models
We will go through each FishMIP model and calculate the mean fish biomass for the decade between 2005 and 2014 (last decade of `historical` period), and for the period between 2041 and 2050 (for the two emission scenarios), and between 2091 and 2100 (two scenarios). Finally, we will calculate the percentage change between these two decades.  
  
```{r, eval = F}
#Looping through each FishMIP model
for(m in members){
  #Load all data available for a single FishMIP model
  df_model <- str_subset(global_files, m) |> 
    map(~readRDS(.)) |> 
    map_df(~bind_rows(.)) 
  
  #Apply MEASO mask
  df_measo <- df_model |> 
    #Apply mask
    left_join(measo_mask, by = join_by(x, y)) |> 
    #Calculate mean per ensemble member
    group_by(year, mem, esm, scenario, id) |> 
    summarise(mean_bio = weighted.mean(biomass, area_m, na.rm = T))
  
  #Apply CCAMLR mask
  df_ccamlr <- df_model |> 
    #Apply mask
    left_join(ccamlr_mask, by = join_by(x, y)) |> 
    #Calculate mean per ensemble member
    group_by(year, mem, esm, scenario, GAR_Short_) |> 
    summarise(mean_bio = weighted.mean(biomass, area_m, na.rm = T))
  
  #Create name to save file  
  f_out_measo <- file.path(out_folder, 
                       str_c(m, "_yearly_perc_bio_change_data_measo.csv"))
  
  f_out_ccamlr <- file.path(out_folder, 
                       str_c(m, "_yearly_perc_bio_change_data_ccamlr.csv"))
  
  #Saving results for each model
  df_measo |> 
    fwrite(f_out_measo)
  
  df_ccamlr |> 
    fwrite(f_out_ccamlr)
}
```
  
## Calculating mean for reference period
Plots created here will show percentage change in biomass over time under two emissions scenarios: `SSP1-2.6` and `SSP5-8.5`. Our reference period is 2005-2014.  
  
### MEASO
  
```{r}
#Listing all relevant files to calculate biomass projections
bio_data_measo <- list.files(out_folder, 
                       pattern = "_yearly_perc_bio_change_data_measo.csv",
                       full.names = T) |> 
  map_df(~fread(.)) |> 
  #Remove rows outside regions of interest
  drop_na(id)

#Calculating reference period
ref_bio_measo <- bio_data_measo |> 
  filter(year >= 2005 & year <= 2014) |> 
  #Calculate mean per ensemble member
  group_by(mem, esm, id) |> 
  summarise(ref_bio_measo = mean(mean_bio, na.rm = T))

bio_data_measo <- bio_data_measo |> 
  #Add reference period to yearly biomass data
  left_join(ref_bio_measo, by = c("mem", "esm", "id")) |> 
  #Calculate percentage change per model/year(scenario)/region
  mutate(perc_change = ((mean_bio-ref_bio_measo)/ref_bio_measo)*100) |> 
  #Calculate mean and standard deviation for percentage change 
  #per year(scenario)/region
  group_by(year, scenario, id) |> 
  summarise(mean_change = mean(perc_change, na.rm = T),
            sd_change = sd(perc_change, na.rm = T)) |> 
  #Add names in FAO-EEZ keys
  left_join(measo_keys, by = "id") |> 
  ungroup()

#Create name to save file  
f_out <- file.path(out_folder, "ensemble_perc_bio_change_data_measo.csv")

#Saving results for each model
bio_data_measo |> 
  write_csv(f_out)
```
  
### CCAMLR
  
```{r}
#Listing all relevant files to calculate biomass projections
bio_data_ccamlr <- list.files(out_folder, 
                       pattern = "_yearly_perc_bio_change_data_ccamlr.csv",
                       full.names = T) |> 
  map_df(~fread(.)) |> 
  #Remove rows outside regions of interest
  drop_na(GAR_Short_)

#Calculating reference period
ref_bio_ccamlr <- bio_data_ccamlr |> 
  filter(year >= 2005 & year <= 2014) |> 
  #Calculate mean per ensemble member
  group_by(mem, esm, GAR_Short_) |> 
  summarise(ref_bio_ccamlr = mean(mean_bio, na.rm = T))

bio_data_ccamlr <- bio_data_ccamlr |> 
  #Add reference period to yearly biomass data
  left_join(ref_bio_ccamlr, by = c("mem", "esm", "GAR_Short_")) |> 
  #Calculate percentage change per model/year(scenario)/region
  mutate(perc_change = ((mean_bio-ref_bio_ccamlr)/ref_bio_ccamlr)*100) |> 
  #Calculate mean and standard deviation for percentage change 
  #per year(scenario)/region
  group_by(year, scenario, GAR_Short_) |> 
  summarise(mean_change = mean(perc_change, na.rm = T),
            sd_change = sd(perc_change, na.rm = T)) |> 
  #Add names in FAO-EEZ keys
  left_join(ccamlr_keys, by = "GAR_Short_") |> 
  ungroup()

#Create name to save file  
f_out <- file.path(out_folder, "ensemble_perc_bio_change_data_ccamlr.csv")

#Saving results for each model
bio_data_ccamlr |> 
  write_csv(f_out)
```
  
## Defining basic plot design
  
```{r}
base_gg <- list(geom_line(linewidth = 0.5),
                #Adding no change line at 0 for reference 
                geom_hline(yintercept = 0, color = "#709fcc", linewidth = 0.65, 
                           linetype = 2),
                #Adding line dividing historical period and future projections
                geom_vline(xintercept = 2015, color = "#709fcc", 
                           linewidth = 0.65),
                #Adding SD as shading 
                geom_ribbon(aes(ymin = mean_change-sd_change,
                                ymax = mean_change+sd_change, fill = scenario),
                            alpha = 0.3, color = NA),
                #Manually setting colours to be used in plots
                scale_color_manual(values = c("historical" = "black",
                                              "ssp126" = "#33bbee", 
                                              "ssp585" = "#ee3377"), 
                                   name = "Scenarios",
                                   labels = c("Historical", "SSP1-2.6", 
                                              "SSP5-8.5")),
                scale_fill_manual(values = c("historical" = "black", 
                                             "ssp126" = "#33bbee", 
                                             "ssp585" = "#ee3377"), 
                                  name = "Scenarios",
                                  labels = c("Historical", "SSP1-2.6", 
                                             "SSP5-8.5")),
                guides(color = guide_legend(nrow = 1, title.position = "left")),
                theme_bw(),
                labs(y = "Change in exploitable\nfish biomass (%)"),
                theme(legend.position = "top", legend.justification = "center", 
                      legend.text = element_text(size = 10),
                      panel.grid.minor.y = element_blank(), 
                      axis.title.x = element_blank(),
                      axis.title.y = element_text(size = 12),
                      axis.text.x = element_text(angle = 45, vjust = 0.765, 
                                                 hjust = 0.65)))
```
  
## Plotting temporal change

### MEASO
```{r}
bio_data_measo |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  facet_wrap(~name, scales = "free_y", ncol = 3)
```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "measo_perc_change.pdf"), 
       device = "pdf", width = 15, height = 12)
```
  
### CCAMLR
```{r}
bio_data_ccamlr |> 
  drop_na() |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  facet_wrap(~GAR_Name, scales = "free_y", ncol = 3)
```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "ccamlr_perc_change.pdf"), 
       device = "pdf", width = 15, height = 15)
```


