---
title: "FishMIP models for the Southern Ocean"
author: "Denisse Fierro Arcos"
format: gfm
toc: true
editor: source
---

# FishMIP regional models for the Southern Ocean
[FishMIP has over 30 regional marine ecosystem models](https://fishmip.org/modellingteams.html), with six models located within the Southern Ocean. Here we will create a map of the Southern Ocean regional models.  
  
## Loading libraries
  
```{r}
library(sf)
library(dplyr)
library(rnaturalearth)
library(ggplot2)
```

  
  
## Loading FishMIP regional models
A shapefile with all FishMIP regional models can be downloaded from our [THREDDS server](http://portal.sf.utas.edu.au/thredds/catalog/gem/fishmip/FishMIP_regions/catalog.html).  
  
```{r}
#Location of regional shapefile
reg_path <- "../../FishMIP_regions/Outputs/FishMIP_regional_models/FishMIP_regional_models.shp"

#Loading shapefile
fishmip_reg <- read_sf(reg_path) |> 
  #Keeping Southern Ocean regional models
  filter(str_detect(region, "Southern Ocean|Prydz|Antarctica|Kerguelen")) |> 
  #South Polar Stereographic projection
  st_transform(crs = 3976)
```

**TO DO:** Correct MICE regional boundaries so they reproject correctly.

```{r}
fishmip_reg |>
  ggplot()+
  geom_sf(aes(fill = region), alpha=0.5)
```

