---
title: "Map of CCAMLR MPA Planning Domains and MEASO regions"
author: "Denisse Fierro Arcos"
format: gfm
toc: true
editor: source
---

# CCAMLR MPA Planning Domains and MEASO
In this notebook, we will create a map with the limits of the nine marine protected areas (MPA) planning domains from the Commission for the Conservation of Antarctic and Marine Living Resources (CCAMLR). These [nine domains](https://www.ccamlr.org/en/science/mpa-planning-domains) are used for planning and reporting on the development of MPAs. The limits of the Marine Ecosystem Assessment for the Southern Ocean (MEASO) are also included in the map. [MEASO](https://sokiaq.atlassian.net/wiki/spaces/MEASO/overview) offers standardised regions that researchers can use to assess and report changes in Southern Ocean ecosystems.  
  
## Loading libraries
  
```{r}
library(sf)
library(dplyr)
library(stringr)
library(rnaturalearth)
library(ggplot2)
library(measoshapes)
```
  
## Loading relevant shapefiles
  
```{r}
#Location of CCAMLR shapefile
ccamlr_path <- "../../SO_shapefiles/CCAMLR_MPA_Planning_Domains/CCAMLR_MPA_Planning_Domains.shp"
#Loading shapefile
ccamlr <- read_sf(ccamlr_path) 

#Loading MEASO regions
measo <- measo_regions05 |> 
  #Remove temperate regions
  filter(str_detect(name, "T$", negate = T)) |> 
  #Matching CCAMLR CRS
  st_transform(st_crs(ccamlr))

#Getting Antarctica boundaries
world <- ne_countries(returnclass = "sf") |> 
  st_transform(crs = st_crs(ccamlr))
```
  
## Plotting Southern Ocean models 
  
```{r}
ggplot()+
  #Plot CCAMLR
  geom_sf(data = ccamlr, aes(fill = Location))+
  #Choosing colourbling friendly palette
  scale_fill_brewer(type = "qual", palette = "Paired")+
  #Overlaying MEASO boundaries as lines
  geom_sf(data = measo, color = "#999999", fill = NA, linewidth = 0.5, 
          linetype = "dashed")+
  #Changing legend title
  guides(fill = guide_legend(title = "CCAMLR MPA domains"))+
  #Add world
  geom_sf(data = world)+
  #Setting map limits
  lims(x = c(-5774572.727594968, 5774572.727594968), 
       y = c(-5774572.727594968, 5774572.727594968))+
  #Applying B&W theme
  theme_bw()
```
  
### Saving plot
  
```{r}
#| eval: false
ggsave("../outputs/ccamlr_measo_map.pdf", device = "pdf", width = 9,
       height = 7)
```

