---
title: "Mapping changes to total consumer biomass in the Southern Ocean"
author: "Denisse Fierro Arcos"
date: "2024-02-19"
format: gfm
toc: true
editor: visual
---

# Changes in total consumer biomass in the Southern Ocean

In this notebook, we will use all FishMIP global models to calculate the mean ensemble percentage change in total consumer biomass (TCB) in the Southern Ocean for the decade ending in 2050. The reference period for this calculation is 2005-2014. We define TCB as biomass for all marine organisms with weights between 10g and 100Kg.  
  
## Loading libraries

```{r}
#| warning: false
#| output: false

#Data wrangling
library(tidyverse)
library(data.table)
# library(janitor)
#Plotting
library(ggnewscale)
#Dealing with raster data
# library(terra)
# library(tidyterra)
#Dealing with vector data
library(sf)
#Base maps
library(rnaturalearth)
#Color palettes
library(cmocean)
#Combining plots
library(cowplot)
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
